\chapter{Introduction}
\setcounter{section}{-1}
\section{Preliminaries}
\subsection*{Modalities}
\begin{itemize}
    \item Oral exams, 10.02. or 24.03.; register online until 12.12. by katrin.kopp@uni-hamburg.de
    \item Office hours on website
    \item Exercises every second week; half of points needed; attendance on 5 or more exercise classes required; exercises have to be presented ; register in Moodle for exercises
\end{itemize}

\subsection*{Motivation}
Given a function \[
f: X \to \mathbb{R}
\] such that $X$ is an (in general, infinite-dimensional) Banach space, we want to minimize $f(x)$ for $x \in X^\text{ad}$, the set of admissible values.
\begin{eg}
    There are plenty of real-world applications:
    \begin{itemize}
        \item modelling of elastic bodies, damage calculation
        \item optimal design (buildings, ships, aircrafts,  \dots )
        \item optimal control (trajectories of spacecrafts, \dots)
        \item inverse problems (seismic tomography, MRT, \dots)
    \end{itemize}
\end{eg}
\begin{theorem}
    Let \[
        f: X^\text{ad} \subseteq \mathbb{R}^n \to \mathbb{R}
    \] be a continuous function and $X^\text{ad}$ be bounded and closed. Then, $\min_{x \in X^\text{ad}} f(x)$ is attained. 
\end{theorem}
\begin{proof}
    Take a minimizing sequence $(x_k)_{k \in \mathbb{N}} \in X^\text{ad}$, i.e. a sequence with \[
        f(x_k) \to \inf_{x \in X^\text{ad}} f(x)
    .\] Since $(x_k)$ is bounded, there exists a convergent subsequence $(x_n) \to x$. Now $x$ is a solution of our problem since:
    \begin{enumerate}
    \item \[f(\lim_{k \to \infty} x_k) = \lim_{k \to \infty} f(x_k) = \inf_{x \in X^\text{ad}} f(x)\] since $f$ is continuous.
    \item Since $X^\text{ad}$ is closed, $x \in X^\text{ad}.$  
    \end{enumerate}
\end{proof}

\begin{eg}
    Consider the optimization problem \[
        \min \int_{-1}^1 (x(s)-x^d(s))^2 \, ds
    \] s.t. $x \in \mathcal{C}([-1,1])$, $-1 \leq x(s) \leq 1 \, \forall s \in [-1,1]$ with \[
    x^d(s) = \begin{cases} -1 &s<0\\ 1 &s \geq 0 \end{cases}
    .\] 
    \begin{enumerate}
        \item \[
            X^\text{ad} = {v \in \mathcal{C}([-1,1]) \mid -1 \leq v(s) \leq 1 \forall s \in [-1,1] }\]
            \item $X^\text{ad}$ is closed: $v_k \to v$ in $\mathcal{C}$, so $v_k(s) \to v(s) \forall s \in [-1,1]$ since $v_k(s) \in [-1,1]$ is closed, so $v(s) \in [-1,1]$.
            \item $f$ is continuous, so if $v_k \to v$ in $\mathcal{C}$, we have
        \begin{align*}
            |f(v_k) - f(v)| &= \int_{-1}^1 (v_k(s)^2 - 2v_k(s)x^d(s)+x^d(s)^2 - (v(s)^2-2v(s)x^d(s)+x^d(s)^2)) \, ds \\
                            &= \left| \int_{-1}^1 v_k(s)^2 - v(s)^2 - 2(v_k(s)-v(s))x^d(s) \, ds \right| \\
                            &= 2 \cdot 2 \max(v_k(s)-v(s)) + 4\|x^d\|_\infty \max |v_k(s)-v(s)| \to 0
        \end{align*}
        . However, there is no solution potential minimizing sequence \[
            x_n(s) = \begin{cases} 1 &s > \frac{1}{n}\\ n \leq &-\frac{1}{n} \leq s \leq \frac{1}{n} \\ -1 &s<\frac{-1}{n} \end{cases}
        .\] 
    \item To see this, consider
        \begin{align*}
            0 &\leq f(x_n) = \int_{- \frac{1}{k}}^0 (ns +1)^2 \, ds + \int_0^{\frac{1}{n}} (ns-1)^2 \,ds \\
              &=2 \cdot \int_0^{\frac{1}{n}} (ks-1)^2 \, ds \leq \frac{2}{n} \to 0
        \end{align*}
        but \[
        f(x) = 0 \iff (x(s)-x^d(s))^2=0 \forall s \implies x=x^d \notin \mathcal{C}
        .\] 
    \end{enumerate}
\end{eg}
It can be even worse:
\begin{eg}
    The Lavrentier phenomenon is concerned with the functional 
    \[
        \mathcal{J}[u]=\int_0^1 (u(t)^3-t)^2 u'(t)^6 \, dt \to \min
    \] s.t. $u(0)=0$ and $u(1)=1$.
    \begin{enumerate}
        \item $u \in \mathcal{C}^1((0,1))$ leads to $v(t)=\sqrt[3]{t}$ with $\mathcal{J}[v]=0$.
        \item $u \in \mathcal{C}^{0,1}((0,1))$ but $\inf \mathcal{J}[v] \geq c >0$.
    \end{enumerate}
\end{eg}
\section{Existence of Solutions}
Consider a problem $(P)$ consisting of finding $\min f(x)$ s.t. $\sigma(x) \in K$ under the following assumptions $(A1)$:
\begin{enumerate}
    \item $f: X \to \mathbb{R}$, $X$ is Banach and $f$ is lower semi-continuous.
    \item $\sigma: X \to Z$, $Z$ is Banach and $\sigma$ is continuous.
    \item $K \subset Z$ closed and convex
\end{enumerate}
\subsection*{Existence by Compactness}
\begin{theorem}[Existence by Compactness]
    Assuming $(A1)$, assume also that there is $x^0 \in X^\text{ad}$ such that \[
        \mathscr{L}_f(x^0) := \{x \in X^\text{ad} \mid f(x) \leq f(x^0)\} 
    \] is compact and non-empty. Then, $(P)$ has a solution. 
\end{theorem}
\begin{proof}
    Let $(x_k)_k \in X^\text{ad}$ be a minimizing sequence such that $f(x_k) \to \inf_{x \in X^\text{ad}} f(x)$. W.l.o.g., we can assume $x_k \in \mathscr{L}_f(x^0)$, so $x_k \to \overline{x} \in \mathscr{L}_f(x^0)$ since we have a convergent subsequence. Hence,
    \[
        f(\overline{x}) = f(\lim x_k) \leq \liminf_{k \to \infty} f(x_k) = \inf_{x \in X^\text{ad}} f(x)
    ,\] and therefore $\overline{x} \in X^\text{ad}$ (lower subcontinuity).
\end{proof}
